### 2025-01-31 Algorithms/water_height.tex
Add pseudo code on water height calculation

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Iterative Retrieval of Water Height}
\label{alg:compute_water_height}
\begin{algorithmic}[1]
\Require A position $\mathbf x$ to evaluate the water height and the horizontal and vertical displacements: $\mathbf{D}(\mathbf x,t)$, $h(\mathbf x,t)$
\State Initialize displacement $\mathbf{W} = (0,0,0)$
\For{$k = 1$ to $N_{\text{iter}}$} \Comment{Iterative correction to account for displacement}
    \State Compute shifted position: $\mathbf{W} = \mathbf{x} - \mathbf{W}$
    \State Compute horizontal displacement at $\mathbf{W}$ 
    \Comment{The sum of bilinear interpolation of the displacement on each cascades}
\EndFor
\State \Return The vertical component of $W$
\end{algorithmic}
\end{algorithm}