### 2025-02-04 Algorithms/fluid_to_solid.tex
Add implementation details about the sum of forces

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Fluid-to-Solid Interaction Algorithm}
\label{alg:fluid_to_solid}
\begin{algorithmic}[1]
\For{all triangles $T_i$} \Comment{Iterate over all triangles}
\State Compute geometrical parameters \Comment{Algorithm \ref{alg:geom_params}}
\State Compute drag forces $\mathbf{F_d}$ for each triangles\Comment{Section \ref{subsubsec:dragForce}}
\EndFor
\State Apply parallel reduction to aggregate forces and geometrical parameters.
\State Transfer reduced forces and parameters to the CPU
\State Compute buoyancy force $\mathbf{F_b}$ from geometrical parameters \Comment{Section \ref{subsubsec:buoyancy}}
\State Integrate total forces and moments in the physics engine
\end{algorithmic}
\end{algorithm}